### 2022-09-15 reports/paper/03-results.rmd
Another draft of  manuscript

Table 1 presents summary statistics of exposure and covariates for the full study population and for those diagnosed with NHL between 1985 and 2004. The cohort is predominantly white (66%) and male (87%). Plant 2 employed the largest number of workers while Plant 1 employed the fewest. The median year of hire among those diagnosed with NHL was 1959 whereas the median year of hire in the full study population was 1967. Age at hire was approximately the same among those with NHL and the full study population. Median lagged cumulative exposure to all three MWF types was higher among NHL cases. Soluble MWFs were the most widely used MWF type, with approximately 90% of workers ever exposed. Median cumulative exposure among exposed was 6.5 times higher for soluble than for straight MWFs. Figure 1 shows median average annual exposure to the three MWF types among exposed workers over calendar time. Exposure to MWF generally followed a downward trend over time. Figure 2 shows the post-intervention distribution of the logarithm of cumulative exposure to soluble MWF at end of follow-up if hypothetical limits on average annual exposure were enforced for all workers starting at hire. The interventions shifted the distribution of cumulative exposure rather than deterministically assigning exposure to a particular value.

@@

![Figure 1. Median average annual exposure to straight, soluble, and synthetic metalworking fluids among exposed workers over time.](`r here::here("resources/images", "exposure.png")`){width=6.5in}

\

![Figure 2. Post-intervention distribution of cumulative exposure to soluble metalworking fluids at end of follow-up under hypothetical limits on average annual exposure (mg/m^3^). The rugplot shows maximum cumulative exposure accrued by non-Hodgkin lymphoma cases.](`r here::here("resources/images", "shift-soluble.png")`){width=4in}

```{r}
intervened_py <- readRDS(here::here("resources", "percent-py-intervened.rds"))
# paste(paste0("\"", intervened_py$dt.name, "\""), collapse = ",\n") %>% clipr::write_clip()
dat.names <- data.table(
	dt.name = c(
		"dat.sol.reduced", "dat.sol.reduced2", "dat.sol.reduced10",
		"dat.sol_total.reduced", "dat.sol_total.reduced2",
		"dat.str.reduced", "dat.str.reduced2", "dat.str.reduced10",
		"dat.str_total.reduced", "dat.str_total.reduced2",
		"dat.syn.reduced", "dat.syn.reduced2", "dat.syn.reduced10",
		"dat.syn_total.reduced", "dat.syn_total.reduced2"),
	variable.dat = c(
		"soluble", "soluble2", "soluble10",
		"total by soluble", "total by soluble2",
		"straight", "straight2", "straight10",
		"total by straight", "total by straight2",
		"synthetic", "synthetic2", "synthetic10",
		"total by synthetic", "total by synthetic2"
	)
)
intervened_py <- as.data.table(merge(
	intervened_py, dat.names,
	by = "dt.name"))
intervened_py <- rbindlist(list(
	data.table(dt.name = NA,
						 mwf = NA, percent_py_intervened = 0,
						 variable.dat = "natural"),
	intervened_py
))
```

@@

# Collect point estimates
point_estimates <- data.table(
	variable = c(
		"straight", "total by straight", "straight2", "total by straight2",
		"straight10",
		"soluble", "total by soluble", "soluble2", "total by soluble2",
		"soluble10",
		"synthetic", "total by synthetic", "synthetic2", "total by synthetic2",
		"synthetic10",
		"scale all", "scale all2", "natural"),
	estimate = c(
		h.reduction.str, h.reduction.str_total, h.reduction.str2, h.reduction.str_total2,
		h.reduction.str10,
		h.reduction.sol, h.reduction.sol_total, h.reduction.sol2, h.reduction.sol_total2,
		h.reduction.sol10,
		h.reduction.syn, h.reduction.syn_total, h.reduction.syn2, h.reduction.syn_total2,
		h.reduction.syn10,
		h.reduction.total, h.reduction.total, h.baseline)
)

@@

```{r}
intervention.labels <- data.table(
	variable = c(
		"natural", "straight", "soluble", "synthetic",
		"straight2", "soluble2", "synthetic2",
		"straight10", "soluble10", "synthetic10",
		"total by straight", "total by soluble", "total by synthetic",
		"total by straight2", "total by soluble2", "total by synthetic2",
		"scale all", "scale all2"
	),
	label = c(
		"None", rep("0.5", 3), rep("0.25", 3), rep("0.05", 3),
		"max(0, 0.5 - sol - syn)", "max(0, 0.5 - str - syn)", "max(0, 0.5 - str - sol)",
		"max(0, 0.25 - sol - syn)", "max(0, 0.25 - str - syn)", "max(0, 0.25 - str - sol)",
		"scale all to 0.5", "scale all to 0.25"
	)
)
# Add intervention labels
rr <- merge(
	rr, intervention.labels[,.(variable = paste0(variable, ".natural"), label)],
	all.x = T,
	by = "variable"
)
```

@@

The observed risk of NHL over the 20-year follow-up period was 6.65 per 1000. Table 2 presents the hazard-extended ICE parametric g-formula estimates of the risk and risk ratios contrasting hypothetical limits on exposure to soluble MWF to no limit, with elimination of competing risks throughout. The estimated risk under no limit on MWF exposure was 9.56 (8.15, 10.89) per 1000. Stronger limits on average annual exposure to soluble MWFs resulted in monotonically stronger reductions in the risk of NHL. Capping average annual exposure to soluble MWFs at the NIOSH REL of 0.5 mg/m^3^ and a tenth of the REL resulted in a risk of 8.30 (6.52, 10.19) and 7.52 (5.73, 9.51) per 1000, respectively. The risk ratios contrasting these hypothetical limits to no limit were 0.87 (0.72, 1.02) and 0.79 (0.62, 0.97). Dynamic reductions in soluble exposure with the aim of limiting total MWF exposure to the REL and half the REL also yielded protective risk ratios: 0.84 (0.69, 0.99) and 0.80 (0.64, 0.98). These dynamic interventions attain risk reductions of similar magnitude as the static interventions enforcing exposure limits at 0.25 and 0.05 mg/m^3^, but do so while intervening on a smaller proportion of person-years.

Results for interventions on straight and synthetic MWF are presented in Tables S1 and S2 in the Appendix. The relative risks contrasting interventions on straight and synthetic MWF exposure to the natural course were much smaller in magnitude than those for soluble; none were statistically significant. Since exposure to soluble MWF was greatest in prevalence and magnitude, limits on soluble MWF would result in interventions on a greater percentage of person-years than limits on straight or synthetic MWF. The intervention capping average annual exposure to soluble MWF at the REL affected 23.8% of the person-years in the follow-up period. Limiting straight and synthetic MWF exposure to the REL affected 3.2% and 1.1% of the person-years under follow-up, respectively.

@@

```{r}
rr[grepl("sol|natural\\.natural", variable),.(
	 	label,
	 	`Person-years intervened (%)` = clean_percent_py(variable),
	 	`Risk per 1000` = clean_risk(variable, scale = 1000), # 34737),
	 	`(risk 95% CI)` = clean_risk.ci(variable, scale = 1000), # 34737),
	 	RR              = sprintf("%.2f", round(estimate, 2)),
	 	`(RR 95% CI)`   = clean_rr.ci(estimate, lower, upper)
	 )][c(1, 2, 4, 3, 5, 6),] -> sol.tab

@@

```
\

```{r}
rr[grepl("str|natural\\.natural", variable),.(
	 	label,
	 	`Person-years intervened (%)` = clean_percent_py(variable),
	 	`Risk per 1000` = clean_risk(variable),
	 	`(risk 95% CI)` = clean_risk.ci(variable),
	 	RR              = sprintf("%.2f", round(estimate, 2)),
	 	`(RR 95% CI)`   = clean_rr.ci(estimate, lower, upper)
	 )][c(1, 2, 4, 3, 5, 6),] -> str.tab

str.tab %>% flextable %>%
	flextable::compose(j = 1, value = as_paragraph(
				"Exposure limit on straight MWF (mg/m", as_sup("3"), ")"),
		part = "header"
	) %>% mwf.flextable %>%
	add_footer_row(values = paste(
		"MWF: metalworking fluid;",
		"sol: exposure to soluble metalworking fluids;",
		"syn: exposure to synthetic metalworking fluids.",
		sep = "\n"), colwidths = 6) %>%
	hline_bottom(part = "footer") %>%
	set_caption(caption = "Counterfactual risks (per 1000)  and risk ratios contrasting interventions on straight MWF to the observed course.") -> str.flextab
saveRDS(str.flextab, "str.flextab.rds")
```
\

```{r}
rr[grepl("syn|natural\\.natural", variable),.(
	 	label,
	 	`Person-years intervened (%)` = clean_percent_py(variable),
	 	`Risk per 1000` = clean_risk(variable),
	 	`(risk 95% CI)` = clean_risk.ci(variable),
	 	RR              = sprintf("%.2f", round(estimate, 2)),
	 	`(RR 95% CI)`   = clean_rr.ci(estimate, lower, upper)
	 )][c(1, 2, 4, 3, 5, 6),] -> syn.tab

>>>>

Table 1 presents summary statistics of exposure and covariates for the full study population and for those diagnosed with NHL between 1985 and 2004. The cohort is predominantly white (66%) and male (87%). The median year of hire among those diagnosed with NHL was 1959 whereas the median year of hire in the full study population was almost a decade later. Age at hire was approximately the same among those with NHL and the full study population. Median lagged cumulative exposure to all three MWF types was higher among NHL cases. Soluble MWFs were the most widely used MWF type, with approximately 90% of workers ever exposed. Median cumulative exposure among the exposed was 6.5 times higher for soluble than for straight MWFs. Figure 2 shows median average annual exposure to the three MWF types among exposed workers over calendar time. Exposure to MWF generally followed a downward trend over time.
<!-- Figure 2 shows the post-intervention distribution of the logarithm of cumulative exposure to soluble MWF at end of follow-up if hypothetical limits on average annual exposure were enforced for all workers starting at hire. The interventions shifted the distribution of cumulative exposure rather than deterministically assigning exposure to a particular value. -->

@@

![Figure 2. Median average annual exposure to soluble, straight, and synthetic metalworking fluids among exposed workers over time.](`r here::here("resources/images", "exposure.png")`){width=6.5in}

```{r}
intervened_py <- readRDS(here::here("resources", "percent-py-intervened.rds"))
# paste(paste0("\"", intervened_py$dt.name, "\""), collapse = ",\n") %>% clipr::write_clip()
dat.names <- data.table(
	dt.name = c(
		"dat.sol.reduced", "dat.sol.reduced2", "dat.sol.reduced10",
		"dat.sol_total.reduced", "dat.sol_total.reduced2",
		"dat.sol_total.reduced10",
		"dat.str.reduced", "dat.str.reduced2", "dat.str.reduced10",
		"dat.str_total.reduced", "dat.str_total.reduced2",
		"dat.str_total.reduced10",
		"dat.syn.reduced", "dat.syn.reduced2", "dat.syn.reduced10",
		"dat.syn_total.reduced", "dat.syn_total.reduced2",
		 "dat.syn_total.reduced10"),
	variable.dat = c(
		"soluble", "soluble2", "soluble10",
		"total by soluble", "total by soluble2",
		 "total by soluble10",
		"straight", "straight2", "straight10",
		"total by straight", "total by straight2",
		 "total by straight10",
		"synthetic", "synthetic2", "synthetic10",
		"total by synthetic", "total by synthetic2",
		 "total by synthetic10"
	)
)
intervened_py <- as.data.table(merge(
	intervened_py, dat.names,
	by = "dt.name"))
intervened_py <- rbindlist(list(
	data.table(dt.name = NA,
						 mwf = NA, percent_py_intervened = 0,
						 variable.dat = "natural"),
	intervened_py
))
```

@@

# Collect point estimates
point_estimates <- data.table(
	variable = c(
		"straight", "total by straight", "straight2", "total by straight2",
		"straight10", "total by straight10",
		"soluble", "total by soluble", "soluble2", "total by soluble2",
		"soluble10", "total by soluble10",
		"synthetic", "total by synthetic", "synthetic2", "total by synthetic2",
		"synthetic10", "total by synthetic10",
		"natural"),
	estimate = c(
		h.reduction.str, h.reduction.str_total, h.reduction.str2, h.reduction.str_total2,
		h.reduction.str10, h.reduction.str_total10,
		h.reduction.sol, h.reduction.sol_total, h.reduction.sol2, h.reduction.sol_total2,
		h.reduction.sol10, h.reduction.sol_total10,
		h.reduction.syn, h.reduction.syn_total, h.reduction.syn2, h.reduction.syn_total2,
		h.reduction.syn10,h.reduction.syn_total10,
		h.baseline)
)

@@

```{r}
intervention.labels <- data.table(
	variable = c(
		"natural", "straight", "soluble", "synthetic",
		"straight2", "soluble2", "synthetic2",
		"straight10", "soluble10", "synthetic10",
		"total by straight", "total by soluble", "total by synthetic",
		"total by straight2", "total by soluble2", "total by synthetic2",
		"total by straight10", "total by soluble10", "total by synthetic10"
	),
	label = c(
		"None", rep("0.5", 3), rep("0.25", 3), rep("0.05", 3),
		"max(0, 0.5 – sol – syn)", "max(0, 0.5 – str – syn)", "max(0, 0.5 – str – sol)",
		"max(0, 0.25 – sol – syn)", "max(0, 0.25 – str – syn)", "max(0, 0.25 – str – sol)",
		"max(0, 0.05 – sol – syn)", "max(0, 0.05 – str – syn)", "max(0, 0.05 – str – sol)"
	)
)
# Add intervention labels
rr <- merge(
	rr, intervention.labels[,.(variable = paste0(variable, ".natural"), label)],
	all.x = T,
	by = "variable"
)
```

@@

The observed risk of NHL over the 20-year follow-up period was 6.65 per 1000. Table 2 presents the hazard-extended ICE parametric g-formula estimates of the risk and risk ratios contrasting hypothetical limits on exposure to soluble MWF to no limit, after elimination of competing risks. The estimated risk under no limit on MWF exposure was 9.56 (8.15, 10.89) per 1000. Stronger limits on average annual exposure to soluble MWFs resulted in monotonically stronger reductions in the risk of NHL. Capping average annual exposure to soluble MWFs at 0.5 mg/m^3^, 0.25 mg/m^3^, and 0.05 mg/m^3^ resulted in a risk of 8.30 (6.52, 10.19), 8.06 (6.15, 10.15), and 7.52 (5.73, 9.51) per 1000, respectively. The risk ratios contrasting these hypothetical limits to no limit were 0.87 (0.72, 1.02), 0.84 (0.68, 1.01), and 0.79 (0.62, 0.97). Dynamic reductions in soluble exposure with the aim of capping total MWF exposure also yielded monotonically protective risk ratios: 0.84 (0.69, 0.99), 0.80 (0.64, 0.98), and 0.72 (0.48, 1.01). The dynamic interventions seeking to cap total MWF exposure at 0.5 mg/m^3^ and 0.25 mg/m^3^ attained risk reductions of similar magnitude as the static interventions enforcing (stronger) exposure limits at 0.25 and 0.05 mg/m^3^, respectively, but do so while intervening on a smaller proportion of person-years.

<!-- Results for interventions on straight and synthetic MWF are presented in Tables S1 and S2 in the Appendix. The relative risks contrasting interventions on straight and synthetic MWF exposure to the natural course were much smaller in magnitude than those for soluble; none were statistically significant. Since exposure to soluble MWF was greatest in prevalence and magnitude, limits on soluble MWF would result in interventions on a greater percentage of person-years than limits on straight or synthetic MWF. The intervention capping average annual exposure to soluble MWF at the REL affected 23.8% of the person-years in the follow-up period. Limiting straight and synthetic MWF exposure to the REL affected 3.2% and 1.1% of the person-years under follow-up, respectively. -->

@@

```{r}
rr[grepl("sol|natural\\.natural", variable),.(
	 	label,
	 	`Person-years intervened (%)` = clean_percent_py(variable),
	 	`Risk per 1000` = clean_risk(variable, scale = 1000), # 34737),
	 	`(risk 95% CI)` = clean_risk.ci(variable, scale = 1000), # 34737),
	 	RR              = sprintf("%.2f", round(estimate, 2)),
	 	`(RR 95% CI)`   = clean_rr.ci(estimate, lower, upper)
	 )][c(1, 2, 4, 3, 5, 7, 6),] -> sol.tab

# saveRDS(sol.tab, here::here('resources', 'sol.tab.rds'))

@@

```

```{r}
rr[grepl("str|natural\\.natural", variable),.(
	 	label,
	 	`Person-years intervened (%)` = clean_percent_py(variable),
	 	`Risk per 1000` = clean_risk(variable),
	 	`(risk 95% CI)` = clean_risk.ci(variable),
	 	RR              = sprintf("%.2f", round(estimate, 2)),
	 	`(RR 95% CI)`   = clean_rr.ci(estimate, lower, upper)
	 )][c(1, 2, 4, 3, 5, 7, 6),] -> str.tab

str.tab %>% flextable %>%
	flextable::compose(j = 1, value = as_paragraph(
				"Exposure limit on straight MWF (mg/m", as_sup("3"), ")"),
		part = "header"
	) %>% mwf.flextable %>%
	add_footer_row(values = paste(
		"MWF: metalworking fluid;",
		"sol: exposure to soluble metalworking fluids;",
		"syn: exposure to synthetic metalworking fluids.",
		sep = "\n"), colwidths = 6) %>%
	hline_bottom(part = "footer") %>%
	set_caption(caption = "Counterfactual risks (per 1000)  and risk ratios contrasting interventions on straight MWF to the observed course.") -> str.flextab
saveRDS(str.flextab, "str.flextab.rds")
```

```{r}
rr[grepl("syn|natural\\.natural", variable),.(
	 	label,
	 	`Person-years intervened (%)` = clean_percent_py(variable),
	 	`Risk per 1000` = clean_risk(variable),
	 	`(risk 95% CI)` = clean_risk.ci(variable),
	 	RR              = sprintf("%.2f", round(estimate, 2)),
	 	`(RR 95% CI)`   = clean_rr.ci(estimate, lower, upper)
	 )][c(1, 2, 4, 3, 5, 7, 6),] -> syn.tab

### 2022-09-15 reports/paper/04-discussion.rmd
Another draft of  manuscript

---
output:
  word_document:
    reference_docx: "~/HeadRs/rmd_templates/style-guide-arial-11.docx"
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
csl: "`r path.expand('~/HeadRs/csl/ama.csl')`"
---

@@

We estimated counterfactual 20-year risks of NHL from 1985 to 2005 in the UAW-GM  Cohort Study under different hypothetical interventions targeting exposure to soluble MWFs using the hazard-extended ICE parametric g-formula. We found a monotonic dose-dependent relationship with lower risk estimates arising from stronger limits. Contrasts in estimated NHL risk were statistically significant under the strongest static intervention, which limited soluble MWF exposure to 0.05 mg/m^3^, and the two dynamic interventions, which reduced exposure so that total MWF exposure would be capped at 0.5 or 0.25 mg/m^3^, if possible. The dynamic interventions, which take into consideration exposure to the other MWF types, may be more efficient at reducing risk of NHL than the static interventions, which ignore exposure to the other MWF types.

Colbeth et al. first reported an association between NHL and MWF exposure in this cohort in their standard survival analyses, which investigated associations between MWF exposure and several cancer incidence outcomes [@Colbeth_2022]. We verified and further examined this association in an approach capable of adjusting for time-varying confounding possibly affected by past exposure. The present approach also yielded marginal rather than conditional estimates of effect while evaluating realistic interventions, and is thus more compatible with a population health framework. Under the assumptions of conditional exchangeability, positivity, consistency, and correct model specification, our estimates are unbiased for the true counterfactual risk under the hypothetical interventions.

Conditional exchangeability means that for all time points, there is no confounding of the relationship between exposure/censoring and NHL status at all future time points given the observed past, including past exposure and covariates. A major threat to conditional exchangeability in longitudinal occupational studies is the HWSE. We limit potential bias due to the HWSE by conditioning on all past cumulative exposure, employment status, and cumulative time off at each time point. Cumulative time off and employment status are reasonable mediators of the causal paths linking past health to future exposure and health, but may not be sufficient to fully adjust for the HWSE. Declines in a worker's health may lead to reductions in work-related exposure without affecting employment status or time off work [@Garcia_2017]. We expect the absence of good time-varying measures of worker health over the life course to result in bias toward the null.

Positivity refers to the need for adequate variation in future exposure among strata formed by observed covariate and intervention-compliant exposure histories. Even under conditional exchangeability, where exposures within these strata may be considered the result of experimental assignment, expected counterfactual outcomes under different exposures may not be estimable if there is sparsity in the observed distribution of exposures [@Maldonado_2002]. We investigated static and dynamic stochastic interventions on soluble MWF exposure which intervened only when average annual exposure exceeded the hypothetical limit under consideration. Hence, our parameters of interest achieve positivity more easily than those for deterministic interventions e.g. setting all to exposed versus unexposed. Nonetheless, violations in positivity were still of concern due to the high dimensionality of covariates, as in common in longitudinal settings. We addressed sparsity by summarizing the 20 years of follow-up over a coarser timescale with only 8 follow-up periods and by representing covariates using fewer categories. Coarsening limits the comprehensiveness of confounding control, but improves positivity. In practice, causal inference using observational data must always balance positivity, covariate adjustment, and model specification [@Petersen_2012]. When estimating causal effects using estimators that do not require exposure modeling, as is the case with g-formula methods, concerns over potential bias due to practical violations in positivity may be relaxed if correct model specification is attained.

The consistency assumption, also known as the no-multiple-versions-of-treatment or stable unit treatment value assumption, is that counterfactual outcomes under each possible exposure value take on a unique value [@Cole_2009;@VanderWeele_2013]. This assumption would be violated if there were multiple versions of treatment causally associated with different outcomes. This basic notion of consistency is violated in our analysis because our exposure of interest is a complex mixture of diverse components with substantial variation over time due to changes in formulation and due to the natural physical, chemical, and biological changes in the MWF over the course of its use [@Howell_2006]. However, causal effects estimates under violations in the consistency assumption are still valid and unbiased if there is adequate adjustment for confounders of the exposure-version relationship [@VanderWeele_2013]. This may be thought of as conditional consistency within strata, in which there is only one version of treatment. Our analysis indexed time periods over calendar time and adjusted for age, year of hire, and plant. In this way, we limited potential for bias due to variation in MWF composition.

Correct model specification is a standard assumption in all analyses. The estimator we applied offers greater statistical efficiency than the classical ICE parametric g-formula estimator because it leverages greater parametric smoothing. In causal analyses of longitudinal cohort studies, both the hazard-extended and classical ICE parametric g-formula estimators are less common than the non-iterative expression (NICE) of the parametric g-formula [@Keil_2014]. However, a major limitation of the NICE g-formula is the g-null paradox: the guaranteed misspecification of parametric models resulting in the false rejection of the null hypothesis when the null is true and there is time-varying confounding affected by past exposure [@Naimi_2015;@McGrath_2022]. As with all ICE g-formula estimators, the estimator we applied is not subject to the g-null paradox. Furthermore, simulation studies show that the variance of the hazard-extended ICE parametric g-formula is similar to that of the NICE parametric g-formula, so we expect the former to be no less conservative than the latter [@Wen_2020].

@@

Much of the existing epidemiologic literature linking occupational and environmental exposures to NHL risk report findings from case-control studies where exposures are measured crudely as binary indicators of exposure or membership in a particular occupational group [@Cano_2001;@Costantini_2001;@Karunanayake_2008;@t-Mannetje_2016]. Associations between occupations and NHL risk vary considerably, but one study of working men in Kansas and Nebraska found strong associations between NHL risk and occupations involving metalworking and motor vehicles [@Zheng_2002]. Both of these occupations may entail exposure to soluble MWFs, which contain a number of additives of concern for human health and NHL risk in particular. Organic compounds containing phosphorous, chlorine, sulfur, nitrogen, and boron are commonly added to soluble MWF to control microbial growth, improve performance under high heat/pressure, and inhibit corrosion [@Evans_2020]. Organophosphorous compounds include organophosphate pesticides, which have been linked to cancer risk in epidemiologic and animal studies. Some were classified as possibly carcinogenic by the IARC [@IARC_2017]. Studies of occupational exposure to chlorinated solvents and pesticides have also been linked to NHL risk [@Cocco_2008;@Purdue_2011;@Cocco_2013;@Vlaanderen_2013;@Callahan_2018]. In 2014, the International Agency for Research on Cancer classified trichloroethylene, tetrachloroethylene, and other chlorinated agents as Group 1 carcinogens [@IARC_2014]. Chlorinated solvents are commonly used as degreasers in industrial settings, but their use in the plants under study here was uncommon and limited to particular operations [@Shrestha_2016]. The structural characteristics shared by MWF additives and known/suspected carcinogens suggest similarities in molecular function and carcinogenesis.

This study investigated the effect of hypothetical limits on average annual exposure to MWF by comparing the standardized distribution of NHL under various distributions of cumulative exposure induced by applying upper bounds to annual exposure. When average annual exposure did not exceed the hypothetical limit, it was left at the observed level of exposure. We selected these hypothetical limits based on the NIOSH REL of 0.5 mg/m^3^ [@niosh_1998]. Without enforcement of the REL, we still observed average annual exposures below the REL for many person-years. If the REL were enforced in the real world, we would not expect reductions in exposure for these low-exposure person-years. Hence, contrasting the counterfactual scenario where all workers experienced annual exposure at the REL to one where all workers experienced annual exposure at some higher level would result in an overestimate of the expected real-world benefit of REL enforcement. By considering stochastic interventions that result in exposure distributions more similar to what we would expect in the real world, our analysis yielded results that more realistically quantify the effects of policy than those investigating deterministic interventions [@Bembom_2007;@Munoz_2012].

@@

Associations between several occupations and risk of NHL have been reported previously, but none evaluated the potential effect of hypothetical limits on occupational exposures [@Ekstrom-Smedby_2006;@Chiu_2015;@t-Mannetje_2016;@Callahan_2018]. We found evidence that exposure to soluble MWF was associated with NHL incidence after adjustment for time-varying confounding affected by prior exposure using the hazard-extended ICE parametric g-formula. Shifting the distribution of cumulative exposure to soluble MWF by limiting average annual exposure resulted in reductions in NHL incidence. Limits on exposure to soluble MWF that take into consideration exposure to straight and synthetic MWFs led to greater reductions NHL incidence while intervening on fewer person-years.

>>>>

---
output:
  word_document:
    reference_docx: "~/HeadRs/rmd_templates/style-guide-arial-11.docx"
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
csl: "`r path.expand('~/HeadRs/csl/ama_2.csl')`"
---

@@

We estimated counterfactual 20-year risks of NHL from 1985 to 2005 in the UAW-GM  Cohort Study under different hypothetical interventions on exposure to soluble MWFs using the hazard-extended ICE parametric g-formula. We found a monotonic exposure-dependent relationship with lower risk estimates arising from stronger limits. Contrasts in estimated NHL risk were statistically significant under the strongest static intervention, which limited soluble MWF exposure to 0.05 mg/m^3^ (a tenth of the current NIOSH REL), and the two dynamic interventions, which sought to limit total MWF exposure to 0.5 and 0.25 mg/m^3^ (the REL and half the REL, respectively) by intervening only on exposure to soluble MWF.

Under the assumptions of conditional exchangeability, positivity, consistency, and correct model specification, our estimates are unbiased for the true counterfactual risk under the hypothetical interventions. Conditional exchangeability means that for all time points, there is no confounding of the relationship between exposure/censoring and NHL status at all future time points given the observed past, including past exposure and covariates. A major threat to conditional exchangeability in longitudinal occupational studies is the HWSE. We limit potential bias due to the HWSE by conditioning on cumulative exposure, employment status, and cumulative time off history at each time point. Cumulative time off and employment status are reasonable mediators of the causal paths linking past health to future exposure and health, but adjustment for these variables may not be sufficient for eliminating bias due to the HWSE. Declines in a worker's health may lead to reductions in work-related exposure without affecting employment status or time off work [@Garcia_2017]. We expect the absence of valid time-varying measures of worker health over the life course to result in bias toward the null.

Positivity refers to the need for adequate variation in future exposure among strata formed by observed covariate and intervention-compliant exposure histories. Even under conditional exchangeability, where exposures within these strata may be considered the result of experimental assignment, expected counterfactual outcomes under different exposures may not be estimable if there is sparsity in the observed distribution of exposures [@Maldonado_2002]. We investigated static and dynamic stochastic interventions on soluble MWF exposure that intervened only when average annual exposure exceeded the hypothetical limit under consideration. Hence, our parameters of interest achieve positivity more easily than those for deterministic interventions e.g. setting all to a single level of exposure. Nonetheless, violations in positivity were still of concern due to the high dimensionality of covariates, as is common in longitudinal settings. We addressed sparsity by summarizing the 20 years of follow-up over a coarser timescale with only 8 follow-up periods and by representing covariates using fewer categories. Coarsening limits the comprehensiveness of confounding control, but improves positivity. In practice, causal inference using observational data must always balance positivity, covariate adjustment, and model specification [@Petersen_2012]. When estimating causal effects using estimators that do not require exposure modeling, as is the case with ICE g-formula methods, concerns over potential bias due to practical violations in positivity may be relaxed if correct model specification, discussed below, is attained.

The consistency assumption, also known as the no-multiple-versions-of-treatment or stable unit treatment value assumption, is that counterfactual outcomes under each possible exposure value take on a unique value [@Cole_2009;@VanderWeele_2013]. This assumption would be violated if there were multiple versions of treatment causally associated with different outcomes. This basic notion of consistency is violated in our analysis because our exposure of interest is a complex mixture of diverse components with substantial variation over time due to changes in formulation as well as the natural physical, chemical, and biological changes in the MWF over the course of its use and reuse [@Howell_2006]. However, causal effect estimates under violations in the consistency assumption are still valid and unbiased if there is adequate adjustment for confounders of the exposure-version relationship [@VanderWeele_2013]. This may be thought of as conditional consistency within strata, in which there is only one version of treatment. Our analysis indexed time periods over calendar time and adjusted for age, year of hire, and plant. In this way, we limited potential for bias due to variation in MWF composition.

Correct model specification is a standard assumption in all parametric analyses. The estimator we applied offers greater statistical efficiency than the classical ICE parametric g-formula estimator because it leverages greater parametric smoothing. In causal analyses of longitudinal cohort studies, both the hazard-extended and classical ICE parametric g-formula estimators are less common than the NICE parametric g-formula [@Keil_2014]. However, a major limitation of the NICE g-formula is the g-null paradox: the guaranteed misspecification of parametric models resulting in the false rejection of the null hypothesis when the null is true and when there is time-varying confounding affected by past exposure [@Naimi_2015;@McGrath_2022]. As with all ICE g-formula estimators, the estimator we applied is not subject to the g-null paradox. Furthermore, simulation studies show that the variance of the hazard-extended ICE parametric g-formula is similar to that of the NICE parametric g-formula, so we expect the former to be no less conservative than the latter [@Wen_2020].

@@

Much of the existing epidemiologic literature linking occupational and environmental exposures to NHL report findings from case-control studies where exposures are measured crudely as binary indicators of exposure or membership in a particular occupational group [@Cano_2001;@Costantini_2001;@Karunanayake_2008;@t-Mannetje_2016]. Associations between occupations and NHL risk vary considerably, but one study of working men in Kansas and Nebraska found strong associations between NHL risk and occupations involving metalworking and motor vehicles [@Zheng_2002]. Both of these occupations may entail exposure to soluble MWFs, which contain a number of additives of concern for human health and for NHL risk in particular. Organic compounds containing phosphorous, chlorine, sulfur, nitrogen, and boron are commonly added to soluble MWF to control microbial growth, improve performance under high heat/pressure, and inhibit corrosion [@Evans_2020]. Organophosphorus compounds include organophosphate pesticides, which have been linked to cancer risk in epidemiologic and animal studies. Some were classified as possibly carcinogenic by the IARC [@IARC_2017]. Studies of occupational exposure to chlorinated solvents and pesticides have also been linked to NHL risk [@Cocco_2008;@Purdue_2011;@Cocco_2013;@Vlaanderen_2013;@Callahan_2018]. In 2014, the IARC classified trichloroethylene, tetrachloroethylene, and other chlorinated agents as Group 1 carcinogens [@IARC_2014]. Chlorinated solvents are commonly used as degreasers in industrial settings, but their use in the plants under study here was uncommon and limited to particular operations [@Shrestha_2016]. The structural characteristics shared by MWF additives and known/suspected carcinogens suggest potential similarities in their behavior in biological systems.

This study investigated the effect of hypothetical limits on MWF exposure by comparing the standardized distributions of NHL under various distributions of cumulative exposure induced by applying upper bounds to average annual exposure to soluble MWF. When average annual exposure did not exceed the hypothetical limit, it was left at the observed level of exposure. We selected these hypothetical limits based on the NIOSH REL of 0.5 mg/m^3^ [@niosh_1998]. In the real world, there is no enforcement of the REL, but we nonetheless observed average annual exposures below the REL in these GM plants for many years. If the REL were enforced in the real world, we would not expect reductions in exposure for these low-exposure person-years. Hence, contrasting the counterfactual scenario where all workers experienced average annual exposure at the REL to one where all workers experienced average annual exposure at some higher level would result in an overestimate of the expected real-world benefit of REL enforcement. Furthermore, since the REL is for MWF exposure of all types, we evaluated dynamic stochastic interventions that better reflect this target policy because they reduce exposure to soluble MWF with the goal of reducing MWF exposure of all types to a certain limit. The interventions we considered result in exposure distributions more similar to what we would expect in the real world; our analysis yielded results that more realistically quantify the effects of policy than those investigating deterministic interventions [@Bembom_2007;@Munoz_2012].

@@

Associations between several occupations and risk of NHL have been reported previously, but none evaluated the potential effect of hypothetical limits on occupational exposures [@Ekstrom-Smedby_2006;@Chiu_2015;@t-Mannetje_2016;@Callahan_2018]. We found evidence that exposure to soluble MWF was associated with NHL incidence after adjustment for time-varying confounding affected by prior exposure using the hazard-extended ICE parametric g-formula. Reducing cumulative exposure to soluble MWF by enforcing hypothetical, but realistic, interventions on average annual exposure would reduce NHL incidence.

### 2022-09-15 reports/paper/05-appendix.rmd
Another draft of  manuscript

```{=tex}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\arraystretch}{1}
```

@@

```{r, tab.cap.pre="Table A"}
str.flextab <- readRDS("str.flextab.rds")
str.flextab
```

```{r}
syn.flextab <- readRDS("syn.flextab.rds")
syn.flextab
```

>>>>

```{=tex}
\renewcommand{\thetable}{A\arabic{table}}
\renewcommand{\thefigure}{A\arabic{figure}}
\renewcommand{\arraystretch}{1}
```

@@

![Oberved and post-intervention distributions of average annual exposure (mg/m^3^). The rugplot shows maximum cumulative exposure accrued by non-Hodgkin lymphoma cases.](`r here::here("resources/images", "shift-avg-soluble.png")`){width=6in}

![Oberved and post-intervention distributions of cumulative exposure to soluble metalworking fluids (mg/m$^3\cdot$years) at end of follow-up. The rugplot shows maximum cumulative exposure accrued by non-Hodgkin lymphoma cases.](`r here::here("resources/images", "shift-cum-soluble.png")`){width=6in}

```{r, tab.cap.pre="Table A", include=F}
str.flextab <- readRDS("str.flextab.rds")
str.flextab
```

```{r, include=F}
syn.flextab <- readRDS("syn.flextab.rds")
syn.flextab
```
